Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2020-04-06 paper/notes.tex
Notes on the modelling of axon stopping.

the term for the divergence of $\mb{J}$ can be written:
%
\begin{equation}
  \label{eq:divJ}
  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right).
\end{equation}
%
In the barrels model, each $\rho_j(\mb{x})$ was assumed static, allowing the
definition of a time-independent $\mb{g}_i(\mb{x})$:
%
\begin{equation}
\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x}).
\end{equation}
%
Here, we represent the time-dependence of the signalling molecule expression
by noting that $\rho$ is now a function of time.  The error which the growth
cones make in sampling $\rho$ and estimating $\nabla\rho_j(\mb{x})$ is
$\xi_\rho(t)$:
%
\begin{equation}
\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x},t)
+ \xi_\rho(t) \big).
\end{equation}
%
One problem with this representation is that the equations represent a
population model. Within one spatial element of the descretized solution of
the differential equations, the error, $\xi_\rho(t)$ can have only one value at
time $t$; this value is applied to all the individual axon growth cones that
are represented by that element. Nonetheless, it's a useful, and fairly simple
way to represent the guidance error in the system. In practice, we ignore the
variation in $\rho_j$ and write
%
\begin{equation}
\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x})
+ \xi_\rho(t) \big).
\end{equation}

@@

Starting from the branching equation again (but dropping the competition term,
$\chi_i$):

@@

The spatial part of the system can be halted if $D\rightarrow
0, \gamma_{i,j}\rightarrow 0$ as
$f(\gamma_{i,1},\gamma_{i,2},...,\gamma{i,M})\rightarrow 1$. That is, if, for
a particular axon, $i$, a sampling of $\rho_j$ is close to $f(\gamma_{i,j})$
for each $j$, then $D$ is reduced to zero, and likewise, the second term also
reduces to zero. $f()$ is probably $1-$ a (possibly noisy) normalized Gaussian
function, or a capped Gaussian (with height capped by a sigmoid?), with mean
$(\gamma_{i,1},\gamma_{i,2})$ and a width based on the differences between
$\gamma_{i\pm 1,j}$. Something like:

@@

$f$ could be thought of as a functional:
$f_i[\gamma_{i,1},\gamma_{i,2}](\mb{x})$ (if that's the right way to write
it...).

Let's try to write f out. See gaussians.m initially.

@@

Reference \cite{goodhill_development_2005} (from which the headings and quotes
below are taken) gives a list of features which a model of retinotectal
development or regeneration should address.

>>>>

the term for the divergence of $\mb{J}$ can be written:
%
\begin{equation}
  \label{eq:divJ}
  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right).
\end{equation}
%
In the barrels model, each $\rho_j(\mb{x})$ was assumed static, allowing the
definition of a time-independent $\mb{g}_i(\mb{x})$:
%
\begin{equation}
\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x}).
\end{equation}
%
Here, we represent the time-dependence of the signalling molecule expression
by noting that $\rho$ is now a function of time.  The error which the growth
cones make in sampling $\rho$ and estimating $\nabla\rho_j(\mb{x})$ is
$\xi_\rho(t)$:
%
\begin{equation}
\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x},t)
+ \xi_\rho(t) \big).
\end{equation}
%
One problem with this representation is that the equations represent a
population model. Within one spatial element of the discretized solution of
the differential equations, the error, $\xi_\rho(t)$ can have only one value at
time $t$; this value is applied to all the individual axon growth cones that
are represented by that element. Nonetheless, it's a useful, and fairly simple
way to represent the guidance error in the system. In practice, we ignore the
variation in $\rho_j$ and write
%
\begin{equation}
\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x})
+ \xi_\rho(t) \big).
\end{equation}

@@

Here, we seek to model the `pure' chemospecificity theory in which axons are
guided by orthogonal morphogen gradients until they reach their destination,
which they must be able to sense. The axons, guided by the \emph{gradient} of
the morphogen concentration are assumed also to be able to sample
the \emph{absolute} concentration of the signalling molecules.  A reduction in
branching, and possibly an enhancement of synaptogenesis are triggered when
the concentrations of the two orthogonal morphogens are simultaneously
measured to be within the axon's `stopping' ranges.

To find a way to implement this in a population model similar to the
competitive model, we start from the branching equation (dropping the
competition term, $\chi_i$):

@@

Spatial transfer of information in the system can be halted if
$D\rightarrow 0$ and $\gamma_{i,j}\rightarrow 0$.
%$f(\gamma_{i,1},\gamma_{i,2},...,\gamma{i,M})\rightarrow 1$.

This can be achieved with a function of the gammas and $\mb{x}$:

@@

Where $f$ is a function which permits branching far away from a given axon's
target region and tends to zero inside the target region. It represents an
axon's internal `stopping process'. Start with a Gaussian function:

\begin{equation}\label{eq:g}
g(\mb{x}) = \mathrm{e}^{-\big(\frac{\mb{x}-\mb{\gamma}}{2\sigma}\big)^2 }
\end{equation}

where $\mb{\gamma}$ gives the location of the peak and $\sigma$ is the width
of the peak. This is squashed through an inverting sigmoid, with sharpness $s$:

\begin{equation}\label{eq:f}
f(\mb{x}) = \frac{2}{1 + \mathrm{e}^{-s g(\mb{x})}} - 1
\end{equation}

In one equation:

\begin{equation}\label{eq:gf}
f(\mb{x}) = \frac{2}{1 + \mathrm{e}^{-s \mathrm{e}^{-\big(\frac{\mb{x}-\mb{\gamma}}{2\sigma}\big)^2 } }} - 1
\end{equation}

@@

Reference \cite{goodhill_development_2005} (from which the headings and quotes
below are taken) gives a list of features which a model of retinotectal
dfevelopment or regeneration should address.
